---
output: latex_fragment
bibliography: ../referencesZotero.bib
csl: ../apa.csl
---

```{r}
#| label: section setup
#| include: false

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "../figures/",
  include = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
#| label: missingness filtering

missInfo <- function(full = NA, reduced = NA){
  missTab <- data.frame(
    nFull = nrow(full),
    nRed = nrow(reduced),
    pptFull = length(unique(full$PID)),
    pptRed = length(unique(reduced$PID)),
    timeFull = length(unique(full$TID)),
    timeRed = length(unique(reduced$TID))
  ) %>%
    mutate(
      nDif = nFull-nRed,
      nDifPerc = nDif/nFull*100,
      pptDif = pptFull-pptRed,
      pptDifPerc = pptDif/pptFull*100,
      timeDif = timeFull-timeRed,
      timeDifPerc = timeDif/timeFull*100
    ) %>%
    select(
      nFull, nRed, nDif, nDifPerc,
      pptFull, pptRed, pptDif, pptDifPerc,
      timeFull, timeRed, timeDif, timeDifPerc
    )
    missTab
}

missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)

missS123 <- rbind(
  missS1,
  missS2,
  missS3
) %>%
  mutate(study = c(1, 2, 3)) %>%
  select(study, everything())

missS123 %>%
  kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      align = "c", #c("l", rep("c", ncol(.)-1)),
      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
      digits=2,
      caption = "Missingness Info by Study") %>%
  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```


Matching these requirements, the data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citep[for the data set see][]{Kreienkamp2022b}. After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days. The short ESM surveys were sent out at around lunch (12pm) and dinner time (7 pm). After the 30-day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire. Participants received either monetary compensation or partial course credits based on the number of surveys they completed. 

The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` ESM measurements. Each of the studies focused on recently arrived first-generation migrants and each study included a number of idiosyncratic variables relevant for the broader research collective. For our empirical example, we focus on the variables that were collected during ESM surveys and were available in all three studies. Variable selection and preparation are described as part of the illustration below but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}. Each study was approved by the ethics board of the university of origin and all participants gave informed consent.

