---
output: latex_fragment
bibliography: ../referencesZotero.bib
csl: ../apa.csl
---

```{r}
#| label: section setup
#| include: false

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "../figures/",
  include = FALSE,
  warning = FALSE,
  message = FALSE
)
```

We first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We see that for some variables the features are generally stronger in separating the clusters. We, for example see that the item on 'how cooperative the interaction was' distinguishes the two clusters across almost all seven features (except for the auto-correlation, see \fgrref[A]{fig:clusterFeatVar}). Compare this to the 'outgroup attitudes' item where the differences between the clusters are much smaller almost all features. We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables. For example, mad and median distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup). These two features stand in stark contrast to other features, such as the lag-1 auto correlations, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This offers some information on which features were are most important in understanding the two extracted groups. Taken these two perspectives together, we can also focus on individual features or variables in particular. We, for example, see a strong difference in the average well-being, where participants in cluster showed a much lower median well-being over the time series. At the same time, in terms of stability, both groups have virtually identical averages MAC statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.

```{=tex}
\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons based on Features and Variables}
  \label{fig:clusterFeatVar}
  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
  \caption*{Note: \\
  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
\end{figure}
```


In the second step, we look at prototypical trajectories of the clusters. For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}. Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}. For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}. Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions. The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup. Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days. This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}). We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).

<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->

Finally, we can also assess the clusters across other person-level variable \citep[e.g.,][]{monden2022}. This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences. To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see $n$ in \fgrref[B]{fig:clusterFeatVar}). While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates are similar for all participants. To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement). When looking at the group comparison, find that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.

In short, we find that that the feature-based clustering discerned two meaningfully different groups of participants. We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attiudes (median) that are also stable over time (MAD, MAC) and tend to increase over the 30 day test period (linear trend). This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (median). This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1). This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (median). This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (median). At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (linear trend), and there was generally less consistency in the experiences they were able to have (MAC, MAD, edf). 

This cluster separation, then, has a number of empirical and practical applications. Firstly, the clusters are descriptive. With tens of variables, hundreds of participants, and thousands of measurements, singular descriptive statistics are often not able to capture the complex patterns that describe the data set. The feature-based clustering offers some direct insight into the complexity within the data set. Secondly, the clusters identify important groups. The adaptive and deteriorating groups offer starting points for empirical exploration as well as practical interventions. Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses. Practitioners, can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face. Thirdly, the feature-based approach is flexible and meaningful. We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.

```{=tex}
\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons over time}
  \label{fig:clusterTs}
  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
  \caption*{Note: \\
  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
\end{figure}
```

```{r}
#| label: average feature by cluster

# CLUSTER VARIABLE LABELS
variableLab <- c(
  "AttitudesDutch" = "Outgroup Attitude",
  "AttitudesPartner" = "Int: Partner Attitude",
  "DaytimeNeedFulfillment" = "Need Fulfillment",
  "exWB" = "Well-Being",
  "InteractionContextAccidental" = "Int: Accidental",
  "InteractionContextCooperative" = "Int: Cooperative",
  "InteractionContextRepresentativeNL" = "Int: Representative",
  "InteractionContextvoluntary" = "Int: Voluntary",
  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
  "KeyNeedFulfillment" = "Int: Need Fulfillment",
  "qualityMeaning" = "Int: Meaningful",
  "qualityOverall" = "Int: Quality"
) %>% enframe(., name = "variable", value = "label")

# ADD VALIDATION VARS:
validationNam <- c(
  "gender.pre",
  "IntGrAnx01.pre",
  "IntGrAnx02.pre",
  "IntGrAnx03.pre",
  "IntGrAnx04.pre",
  "IntGrAnx05.pre",
  "IntGrAnx06.pre",
  "IntGrAnx07R.pre",
  "IntGrAnx08R.pre",
  "IntGrAnx09R.pre",
  "IntGrAnx10R.pre",
  "IntGrAnx11R.pre",
  "IntGrAnx12R.pre",
  "SWL01.pre",
  "SWL02.pre",
  "SWL03.pre",
  "SWL04.pre",
  "SWL05.pre",
  "EvDayDiscr01.post",
  "EvDayDiscr02.post",
  "EvDayDiscr03.post",
  "EvDayDiscr04.post",
  "EvDayDiscr05.post",
  "EvDayDiscr06.post",
  "EvDayDiscr07.post",
  "EvDayDiscr08.post",
  "EvDayDiscr09.post"
)

# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
validation <- rbind(
  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
) %>% 
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  group_by(study, ID) %>%
  summarise(
    gender = gender.pre,
    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
  ) %>%
  ungroup %>%
  distinct %>% 
  mutate(
    gender.z = scale(gender, scale = TRUE),
    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
    SWL.pre.z = scale(SWL.pre, scale = TRUE),
    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
  ) %>%
  select(
    ID,
    #gender.z,
    #IntGrAnx.pre.z,
    #SWL.pre.z,
    EvDayDiscr.post.z
  )
val <- validation %>% select(-ID) %>% names
validationFeature <- validation
for(feature in variableLab$variable) {
  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
}
validationFeature <- validationFeature %>% select(-any_of(val))


# CLUSTER VARIABLE LABELS
variableLab <- c(
  "AttitudesDutch" = "Outgroup Attitude",
  "AttitudesPartner" = "Int: Partner Attitude",
  "DaytimeNeedFulfillment" = "Need Fulfillment",
  "exWB" = "Well-Being",
  "InteractionContextAccidental" = "Int: Accidental",
  "InteractionContextCooperative" = "Int: Cooperative",
  "InteractionContextRepresentativeNL" = "Int: Representative",
  "InteractionContextvoluntary" = "Int: Voluntary",
  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
  "KeyNeedFulfillment" = "Int: Need Fulfillment",
  "qualityMeaning" = "Int: Meaningful",
  "qualityOverall" = "Int: Quality"
) %>% enframe(., name = "variable", value = "label")
validationLab <- c(
  "gender.z" = "Val: Gender",
  "SWL.pre.z" = "Val: Sat. with life",
  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
  "EvDayDiscr.post.z" = " Discrimination"
) %>% enframe(., name = "variable", value = "label")

featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
  merge(., z_data) %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
  ungroup %>% 
  select(-starts_with("ID")) %>%
  pivot_longer(
    cols = -c(cluster),
    names_to = c("variable", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>% 
  mutate(
    nam = variable,
    variable = gsub("\\_.*", "", nam),
    feature = gsub("^.*?\\_", "", nam)
  ) %>% 
  merge(., variableLab) %>%
  select(
    cluster,
    nam,
    variable,
    label,
    feature,
    everything()
  )

featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
  merge(., z_data) %>%
  merge(., validationFeature) %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
  ungroup %>% 
  select(-starts_with("ID")) %>%
  pivot_longer(
    cols = -c(cluster),
    names_to = c("variable", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>% 
  mutate(
    nam = variable,
    variable = gsub("\\_.*", "", nam),
    feature = gsub("^.*?\\_", "", nam)
  ) %>% 
  merge(., variableLab) %>% 
  mutate(feature = stri_replace_all_regex(
    feature,
    pattern=validationLab$variable,
    replacement=validationLab$label,
    vectorize=FALSE
  )) %>% 
  select(
    cluster,
    nam,
    variable,
    label,
    feature,
    everything()
  )

rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
  merge(., raw_data) %>%
  select(
    ID,
    TIDnum,
    cluster,
    any_of(varNamS123)
  ) %>%
  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
  merge(., variableLab)
```

```{r}
#| label: clusterVariableGrid

clusterVariableGrid <- featCluster %>% 
  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  coord_flip() +
  labs(
    x = "Features",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ label) +
  theme_Publication()
clusterVariableGrid
```

```{r}
#| label: clusterFeatureGrid

clusterFeatureGrid <- featClusterVal %>% 
  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  coord_flip() +
  labs(
    x = "Features",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
  theme_Publication()
clusterFeatureGrid
```

```{r}
#| label: clusterFeatVarComb

clusterFeatVarComb <- ggpubr::ggarrange(
  clusterVariableGrid + theme(axis.title.x = element_blank()),
  clusterFeatureGrid,
  ncol = 1,
  labels = c("(A) Variable Focus", "(B) Feature Focus"),
  common.legend = TRUE,
  legend = "bottom",
  align = "v"
) 
ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
```

```{r clusterTS, include=FALSE}
clusterTS <- rawCluster %>% 
  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
  stat_summary(fun=mean, geom="line") +
  facet_wrap(~ label, scales = "free_y") +
  labs(
    x = "Time ID",
    y = "Response (Mean Cluster Response Bold)",
    color = "Cluster"
  ) +
  theme_Publication() 
clusterTS
```

```{r clusterTSTrend, include=FALSE}
clusterTSTrend <- rawCluster %>% 
  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
  geom_smooth(aes(group=cluster)) +
  #stat_summary(fun=mean, geom="line") +
  facet_wrap(~ label, scales = "free_y") +
  labs(
    x = "Time ID",
    y = "Response (Mean Cluster Response Bold)",
    color = "Cluster"
  ) +
  theme_Publication()
clusterTSTrend
```

```{r clusterTsComb, include=FALSE}
clusterTsComb <- ggpubr::ggarrange(
  clusterTS + theme(axis.title.x = element_blank()),
  clusterTSTrend,
  ncol = 1,
  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
  common.legend = TRUE,
  legend = "bottom"
)
ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
```